\Large Last Writer Slices and Communication Traps

### 2013-11-07 paper.tex
committing an update that includes the debugging of apache-2.0.48 memcache bug (21285)

\subsection{Formalism and Correctness}
\label{sec:ctsoundness}
To discuss the correctness of \ctraps, we first formalize trap delivery by
building on our last writer slicing formalism from
Section~\ref{sec:lastwriterslices}.

Given a program $P'$ already instrumented with \lwt updates, the \ctraps
compiler produces $P''$. A thread $t$ in $P''$ executes a modified instruction
sequence, $P''_{t} = \{ T(p^{t}_{j}), I(p^{t}_{1}), p^{t}_{1},$ $\ldots,
T(p^{t}_{j}), I(p^{t}_{k_{t}}), p^{t}_{k_{t}} \}$, where $T(p^{t}_{j})$ is a
trap hook call.  We consider a point $Q = \{q_1, \ldots, q_{e-2}\}$ after $e-2$
instructions of an execution have executed and $q_{e} = p^{t}_{j}$ is about to
execute after its trap hook $T(p^{t}_{j})$ and \lwt update $I(p^{t}_{j})$
execute.  The trap hook $T(p^{t}_{j})$ {\em delivers} a trap before if $LWT[m]
= (t',p^{t'}_{\ell w})$ and $t \ne t'$.  Informally, the trap hook delivers a
trap before the instruction if the thread that last wrote $m$ is different from
the one executing the instruction.  When a trap hook, $T(p^{t}_{j})$, delivers
a trap, thread $t$ executes an arbitrary trap handler $H = \{h^{t}_{1}, \ldots,
h^{t}_{h_{n}}\}$.  Thus, after the trap, the \lwt update, and the instruction
execute, the execution reaches the new point $Q = (q_{1}, \ldots, T(q_{e}),
h^{t}_{1}, \ldots, h^{t}_{h_{n}}, I(q_{e}), q_{e})$, and continues.

A \ctraps implementation must be sound and complete to be correct.  A design is
{\em sound} if it only delivers a trap when threads have communicated.  A
design is {\em complete} if it always delivers a trap when threads have
communicated.  The \ctraps compiler puts a trap hook on all read and write
operations, and only on those operations.  Assuming a correct \lwt, \ctraps is
correct.  Our \ctraps implementation also does not insert any synchronization
into the program, so it comes with the same caveats related to data-race
freedom as the \lwt.  The same arguments for \lwt correctness apply to the
correctness of trap delivery and handling.  Most importantly, data-race freedom
implies that trap hooks, \lwt updates, and accesses are all atomic.

@@

>>>>

%MAYBE PUT THIS BACK IN
%\subsection{Formalism and Correctness}
%\label{sec:ctsoundness}
%To discuss the correctness of \ctraps, we first formalize trap delivery by
%building on our last writer slicing formalism from
%Section~\ref{sec:lastwriterslices}.  
%
%Given a program $P'$ already instrumented with \lwt updates, the \ctraps
%compiler produces $P''$. A thread $t$ in $P''$ executes a modified instruction
%sequence, $P''_{t} = \{ T(p^{t}_{j}), I(p^{t}_{1}), p^{t}_{1},$ $\ldots,
%T(p^{t}_{j}), I(p^{t}_{k_{t}}), p^{t}_{k_{t}} \}$, where $T(p^{t}_{j})$ is a
%trap hook call.  We consider a point $Q = \{q_1, \ldots, q_{e-2}\}$ after $e-2$
%instructions of an execution have executed and $q_{e} = p^{t}_{j}$ is about to
%execute after its trap hook $T(p^{t}_{j})$ and \lwt update $I(p^{t}_{j})$
%execute.  The trap hook $T(p^{t}_{j})$ {\em delivers} a trap before if $LWT[m]
%= (t',p^{t'}_{\ell w})$ and $t \ne t'$.  Informally, the trap hook delivers a
%trap before the instruction if the thread that last wrote $m$ is different from
%the one executing the instruction.  When a trap hook, $T(p^{t}_{j})$, delivers
%a trap, thread $t$ executes an arbitrary trap handler $H = \{h^{t}_{1}, \ldots,
%h^{t}_{h_{n}}\}$.  Thus, after the trap, the \lwt update, and the instruction
%execute, the execution reaches the new point $Q = (q_{1}, \ldots, T(q_{e}),
%h^{t}_{1}, \ldots, h^{t}_{h_{n}}, I(q_{e}), q_{e})$, and continues.
%
%A \ctraps implementation must be sound and complete to be correct.  A design is
%{\em sound} if it only delivers a trap when threads have communicated.  A
%design is {\em complete} if it always delivers a trap when threads have
%communicated.  The \ctraps compiler puts a trap hook on all read and write
%operations, and only on those operations.  Assuming a correct \lwt, \ctraps is
%correct.  Our \ctraps implementation also does not insert any synchronization
%into the program, so it comes with the same caveats related to data-race
%freedom as the \lwt.  The same arguments for \lwt correctness apply to the
%correctness of trap delivery and handling.  Most importantly, data-race freedom
%implies that trap hooks, \lwt updates, and accesses are all atomic.

@@

\paragraph{Apache}
%http://issues.apache.org/bugzilla/show_bug.cgi?id=21287
%concurrency-bugs
We conducted a case study using Apache httpd-2.0.48, which has a bug in its
script cache. The cache uses reference counting to determine when to deallocate
cached objects.  When a thread sees a reference count hit zero, it frees the
object.  The bug is that the check that the count is zero and the deallocation
are not atomic but it should be.  Multiple threads may see the zero reference
count and free the object more than once, leading to a crash.

We debugged this issue by examining the core dump from a crashing execution.
The core dump showed that the memory allocator aborted due to one thread
freeing already freed memory at {\tt mod\_mem\_cache.c}, line 287.  We examined
the last writer table entry for the object being freed at that point.  That
entry revealed that a different thread last wrote that object at the same point
in the code.  The information in the last writer table illustrates why the
double free bug is happening -- two threads end up in the object deallocation
routine for the same object, which should not happen.
